R Notebook

### 2020-05-27 src/testing_file.Rmd
review edit fiddling

```{r}
library("cluster")
pam.res <- pam(pres, 6)
# Visualize
fviz_cluster(pam.res)
```

@@

>>>>

@@

#table 2 made as a plot

gather all the percentages in a single file
```{r}
flst <- list.files("./output/maxent/paper_plots/prob_by_depth/", pattern = "_cellprob_totals.csv", full.names = TRUE)
ptables <- lapply(flst, read.csv, header = TRUE)
pall <- do.call(rbind, ptables)
pall$month <- capitalize(pall$month)

ggplot(data = pall, aes(x = month, y = percent, fill = Category)) + geom_bar(stat = "identity")
```

# find mean bottom depth of study region

```{r}
bathyrast <- raster("./data/env/bathymetric_glorys.asc") 
plot(bathyrast)
```
need to reclassify cells outside the mcp to NA 
```{r}
outsidemcp <- shapefile("./data/env/landp/outside-mcp-wgs84.shp") 
ascdirs <- list.dirs("./data/env/ascii", full.names = FALSE, recursive = FALSE)

ras<- bathyrast
ras[outsidemcp] <- 500000
ras[ras == 500000] <- NA
writeRaster(ras, "./data/env/bathy_trimmed.asc", overwrite=TRUE)
cellStats(ras, stat = "mean", na.rm = TRUE) #1046.574
summary(ras) #313.5291
```

find the number of folds in each month with a TSS <0.6
```{r}
flst <- list.files("./output/maxent/models/test-comb", pattern = "avgAIC.csv", full.names = TRUE, recursive = FALSE)
df <- data.frame(month=character(10), nfolds=numeric(10), stringsAsFactors = FALSE)

for (f in 1:length(flst)){
  af <- read.csv(flst[f])
  af <- head(af, -1)
  mth <- (sapply(strsplit(flst[f], "/"), "[[", 6)) 
  df$month[f] <- stringr::str_sub(mth, end = 3)
  df$'x<0.6'[f] <- sum(af$tss <= 0.6)
  df$nfolds[f] <- nrow(af)
}
write.csv(df, "./output/maxent/models/test-comb/tssfoldcount.csv", row.names = FALSE)
```

boxplots of depth of samples relatvie to the bottom depth
```{r}
presback <- read.csv("./output/bio/presab_nafo_yr.csv", header = TRUE) 
pres <- subset(presback, occurrence == 1)
pres <- subset(pres, select = c(month, bottom_depth_glorys, depth_layer))
pres$depthdiff <- pres$bottom_depth_glorys - pres$depth_layer
pres <- subset(pres, month >2)
pres$month <- as.factor(pres$month)

ggplot(pres, aes(x = month, y = depth_layer)) + geom_boxplot() + scale_y_reverse() + labs(y = "sampling depth")

ggplot(pres, aes(x = month, y = depthdiff)) + geom_boxplot() + labs(y = "sampling depth relative to sea floor (0)")

#historgams of probability by depth

This is figure 4

clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
cumst <- data.frame(0, 0)
colnames(cumst) <- c("depth", "no_cells")
for (cum in 1:length(clogasc_list)){
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cumst[cum, 1] <- str_sub(ascname, 13)
  #cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
  ccnt <- freq(cumasc, useNA='no')
  ccnt <- sum(ccnt[ ,2])
  cumst[cum, 2] <- ccnt
}

write.csv(cumst1, "../output/maxent/paper_plots/cellsperlayer.csv", row.names = FALSE)

get data to plot probability by depth... 
```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
binlst <- c(1:10)

for (m in 1:length(mthord)){
  for (b in 1:length(binlst)){
    mth <- mthord[m]
    bin <- binlst[b]
    # mth <- "dec"
    # bin <- 1
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)

  binaryavgstack <- stack(ascdepth)
  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
  stackstathigh[ ,1] <- c(1:47)
  colnames(stackstathigh) <- c("asclayer", "prob_bin")

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 77, -5)
    dlay <- binaryavgstack[[lyr]]
    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

  }
  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
  }
}

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mth <- mthord[m]
  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
  a <- read.csv(probblst[1])
  b <- read.csv(probblst[2])
  c <- read.csv(probblst[3])
  d <- read.csv(probblst[4])
  e <- read.csv(probblst[5])
  f <- read.csv(probblst[6])
  g <- read.csv(probblst[7])
  h <- read.csv(probblst[8])
  i <- read.csv(probblst[9])
  j <- read.csv(probblst[10])

  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
}

Get proprotions
```{r}
mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
    # d1 <- str_sub(a$Depth[r], end = 4)
    d1 <- a$Depth[r]
	  for (ro in 1:nrow(cumst1)){
      #d2 <- str_sub(cumst1$depth[ro], end = 4)
	    d2 <- cumst1$depth[ro]
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
  }
    a$Depth <- as.numeric(a$Depth)

  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), row.names = FALSE)
}

plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)

313.5291
```{r}
require(data.table)
require(purrr)
require(lubridate)
require(tidyverse)
require(viridis)

filenames <- list.files("./output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
filenames <- filenames[ !grepl("maxkap", filenames) ] #remove all the maxkap files

all <- map_dfr(filenames, function(x) fread(x) %>% 
                       mutate(Month=substr(basename(x),1,3))) %>% 
  mutate(Probability = gsub(">0.9", "0.9 - 1", Probability),
         Depth = factor(round(Depth,1)),
         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
         Month=factor(months(parse_date_time(Month,"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))))
#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
dlst <- unique(all$Depth)
dlst <- as.character(dlst)
dlst <- as.numeric(dlst)
dlst <- sort(dlst)
#make the plot
  ggplot(all) +
  geom_col(aes(Depth, p_cells, fill = Probability),width=1) + 
  coord_flip()+
  facet_wrap(~Month,ncol=2)+
  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1, guide=guide_legend(nrow=2, byrow = TRUE)) +
  theme_classic()+
  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47])) + geom_vline(xintercept=12.5, linetype = 2, col = "light grey") # this is approximate

ggsave("./output/maxent/paper_plots/prob_by_depth_line.png", width = 210, height = 297, units = "mm", dpi = 600)
```

get data to plot maxkap binary by depth plot (Figure 6)

Get proportions...

plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
This is Figure 6

```{r maxkappa habitat by depth}

filenames <- list.files("./output/maxent/paper_plots/prob_by_depth/", pattern = "_maxkap_dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)

#make the plot
  ggplot(all) +
  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
  coord_flip()+
  facet_wrap(~Month,ncol=2)+
  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'), name = "Presence")+
  theme_classic()+
  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47])) +
    geom_vline(xintercept=12.5, linetype = 2, col = "light grey") # this is approximate

ggsave("./output/maxent/paper_plots/maxkap_prob_by_depth_line.png", width = 210, height = 297, units = "mm", dpi = 600)
```

This is Figure 3

Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)

have moved sample files into a separate folder
march,  june, sep, dec
11.7737, 61.1128, 199.79. 628.026

s <- st_read("./data/env/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj)

mcp <- st_read("./output/bio/minimum_convex_poly/mcp-wgs84.shp")

filelist <- list.files("./data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)

x <- stack(file.path("./data/plot2-probmaps/",filelist)) %>% 
  mask(mcp)

df$month <- str_sub(df$variable, start = 16, end = 18)

CapStr <- function(y) {
  c <- strsplit(y, " ")[[1]]
  paste(toupper(substring(c, 1,1)), substring(c, 2),
      sep="", collapse=" ")
}
df$month <- sapply(df$month, CapStr)
df$month_f = factor(df$month, levels=c('Mar','Jun','Sep','Dec'))

ggplot(df)+
  geom_sf(data=mcp,fill='grey')+
  geom_raster(aes(x,y,fill=value))+
  geom_sf(data=mcp,fill="transparent")+
  scale_fill_viridis(option="magma",
                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
  facet_grid(depth~month_f,
             labeller = labeller(depth=depthlab))+
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "right",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
        strip.text.x = element_text(face = "bold", size = 12),
        strip.text.y = element_text(face = "bold", size = 12))

ggsave("./output/maxent/paper_plots/prob-overview_swap.png", width = 210, height = 297, units = "mm", dpi = 600)
```
